---
title: "Homework4"
author: "KiseokUChicago"
date: "2021-02-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=9, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

### Homework4 (for Chapter 5)
## Coding assignment for ECEV 42900
Professor: **Sarah Cobey**, **Greg Dwyer** \
Student: **Kiseok Lee**  

### Exercise 1. Approximating epidemic sizes using Newton’s method.
(1) Use the templates to plot the fraction of recovered population R(infinity) as a function of R0. Values for R0 change from disease to disease and from host population to host population. 
For example, estimates for influenza are R0 = [1.2, 3]; for rubella,
we have R0 = [6, 7]; and for measles, R0 = [16, 18]. 
(2) What fraction of a completely susceptible population will be infected
by measles? rubella? influenza? Change the code above to compute the values.
(3) Please also plot the herd immunity threshold on the same graph. The curves differ. 
(4) Please explain the implications for managing pandemics.

```{r}
## This is for SIR model
Funct<-function(x,R0){
	return(1-exp(-R0*x))
}
DeriFunct<-function(x,R0){
	return(R0*exp(-R0*x))
}

## Prototype example of the method
xn<-1000000 # a very large number
xnp1<-1 # a guesstimate of the root
Eps<-10^(-3) # tolerance
R0 <- 0.5
print(abs(xn-xnp1)>Eps)
while(abs(xn-xnp1)>Eps){
	xn<-xnp1
	xnp1<-xn-Funct(xn,R0)/DeriFunct(xn,R0)
	print(abs(xn-xnp1)>Eps)
	print(xnp1)
}


## Implementation to plot fraction recovered
RInf<-function(R0,Funct,DeriFunct){
	xn<-1000000 # a very large number
	xnp1<-1 # a guesstimate of the root
	Eps<-10^{-3} # tolerance
	# RO <- 0.5
	while(abs(xn-xnp1) > Eps){
		xn<-xnp1
		xnp1<-xn - (Funct(xn,R0)/DeriFunct(xn,R0))
		print(paste0('abs(xn-xnp1) ',abs(xn-xnp1)))
		print(paste0('T or F is ',abs(xn-xnp1) > Eps))
		print(paste0('xn is ',xn))
	  print(paste0('xnp1 is ',xnp1))
	}

	return(xnp1)
}

R0s<-seq(0.01,5,length.out=500)
RecFracs<-rep(0,500)
for (i in 1:length(R0s)){
  print(i)
  print(paste0('RO is ',R0s[i]))
  RecFracs[i]<-RInf(R0s[i],Funct,DeriFunct)
}
plot(RecFracs~R0s,t="l",xlab="R0",ylab="Fraction Recovered")
abline(h=1,col="red")


```





