---
title: "Homework4"
author: "KiseokUChicago"
date: "2021-02-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=9, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

### Homework4 (for Chapter 5)
## Coding assignment for ECEV 42900
Professor: **Sarah Cobey**, **Greg Dwyer** \
Student: **Kiseok Lee**  

### Exercise 1. Approximating epidemic sizes using Newton’s method.
(1) Use the templates to plot the fraction of recovered population R(infinity) as a function of R0. Values for R0 change from disease to disease and from host population to host population. 
For example, estimates for influenza are R0 = [1.2, 3]; for rubella,
we have R0 = [6, 7]; and for measles, R0 = [16, 18]. 
(2) What fraction of a completely susceptible population will be infected
by measles? rubella? influenza? Change the code above to compute the values.
(3) Please also plot the herd immunity threshold on the same graph. The curves differ. 
(4) Please explain the implications for managing pandemics.

```{r}
## This is for SIR model
Funct<-function(x,R0){
	return(1-exp(-R0*x))
}
DeriFunct<-function(x,R0){
	return(R0*exp(-R0*x))
}

## Prototype example of the method
xn<-1000000 # a very large number
xnp1<-1 # a guesstimate of the root
Eps<-10^(-3) # tolerance
R0 <- 0.5
print(abs(xn-xnp1)>Eps)
while(abs(xn-xnp1)>Eps){
	xn<-xnp1
	xnp1<-xn-Funct(xn,R0)/DeriFunct(xn,R0)
	print(abs(xn-xnp1)>Eps)
	print(xnp1)
}


## Implementation to plot fraction recovered
RInf<-function(R0,Funct,DeriFunct){
	xn<-1000000 # a very large number
	xnp1<-1 # a guesstimate of the root
	Eps<-10^{-3} # tolerance
	# RO <- 0.5
	while(abs(xn-xnp1) > Eps){
		xn<-xnp1
		xnp1<-xn - (Funct(xn,R0)/DeriFunct(xn,R0))
		# printing output
	# 	print(paste0('abs(xn-xnp1) ',abs(xn-xnp1)))
	# 	print(paste0('T or F is ',abs(xn-xnp1) > Eps))
	# 	print(paste0('xn is ',xn))
	#   print(paste0('xnp1 is ',xnp1))
	}
	return(xnp1)
}

R0s<-seq(0.01,5,length.out=500)
RecFracs<-rep(0,500)
for (i in 1:length(R0s)){
  # print(i)
  # print(paste0('RO is ',R0s[i]))
  RecFracs[i]<-RInf(R0s[i],Funct,DeriFunct)
}
plot(RecFracs~R0s,t="l",xlab="R0",ylab="Fraction Recovered")
abline(h=1,col="red")


```

### Excercise 2. Visualizing epidemic dynamics
For measles, it has been estimated that # ⇡ 0.2/day and that R0 ⇡ 18.
Use the Runge-Kutta 4th-order ODE solver from the R package
deSolve to integrate the equation above for the disease in a
population for which S(0) = 0.9999, R(0) = 0, I(0) = 0.0001.
Run the model for 30 days with N = 100, 000. How many people
have recovered after 5 days? How many after 10? Plot the
curves of S, I, and R. Gain experience points by experimenting
with di↵erent solvers to see if they have an e↵ect.

#### Plotting SIR curves for 30 days
```{r}
library(deSolve)
library(dplyr)

## the derivative of the logistic
SIR_R<-function(t, state, parameters) { 
  X <- state[1] ## the first element is the density of X at time t
  r <- parameters[1] ## the first parameter is the recovery rate
  R0 <- parameters[2] ## R0
  S0 <- parameters[3] ## S0
  dR<- r*(1- S0*exp(-R0*X) - X)
  return(list(dR)) ## for some reason, you have to return a list
}

SIR_S<-function(t, state, parameters) { 
  X <- state[1] ## the first element is the density of X at time t
  r <- parameters[1] ## the first parameter is the recovery rate
  R0 <- parameters[2] ## R0
  S0 <- parameters[3] ## S0
  dS<- -r*R0*X*(1-X-(1/R0)*(log(S0)-log(X)))
  return(list(dS)) ## for some reason, you have to return a list
}

t  <- seq(0,30,0.5)
r <- 0.2; R0 <- 18; S0 <- 0.9999;
parameters <- c(r = r, R0 = R0, S0 = S0)
x <- c(R = 0)
y <- c(S = 0.9999)

## reasonable numerical solution with rk4 (Runge-Kutta 4th-order ODE solver)
out_R <- as.data.frame(rk4(x, t, SIR_R, parameters))
out_S <- as.data.frame(rk4(y, t, SIR_S, parameters)) ; out_S$S[is.nan(out_S$S)]<-0  # replace NaN with 0
out_I <- left_join(out_R,out_S, by=c('time'='time')) %>% mutate(I=1-R-S)
#

# plot SIR graphs (time step is adjusted to 0.5)
# plot R
plot(out_R$t, out_R$R, type='l', col = "blue", cex = 0.5,  ylim=c(0,1), xlab='time', ylab='fraction of total population', main = 'Plotting S, I, R curves')
# plot S
points(out_S$t, out_S$S, type='l', col = "black", cex = 0.5)
# plot I
points(out_I$t, out_I$I, type='l', col = "red", cex = 0.5)
# legend
legend(27,0.9,legend=c("R", "S","I"),
       col=c("blue", "black",'red'), lty=1, cex=0.8)
  
```

#### Converting y axis into number of people 
Run the model for 30 days with N = 100, 000. How many people
have recovered after 5 days? How many after 10? Plot the
curves of S, I, and R. 

```{r}
## multiplying number of total population
N <- 100000
out_SIR <- left_join(out_R,out_S, by=c('time'='time')) %>% mutate(I=1-R-S) %>% 
  mutate(n_R=R*N) %>% mutate(n_S=S*N) %>% mutate(n_I=I*N)

# plot SIR graphs (time step is adjusted to 0.5)
# plot R
plot(out_SIR$t, out_SIR$n_R, type='l', col = "blue", cex = 0.5, ylim=c(0,N), xlab='time', ylab='# of people', main = 'Plotting SIR curves with number of people')
# plot S
points(out_SIR$t, out_SIR$n_S, type='l', col = "black", cex = 0.5)
# plot I
points(out_SIR$t, out_SIR$n_I, type='l', col = "red", cex = 0.5)
# legend
legend(27,N-10000,legend=c("R", "S","I"),
       col=c("blue", "black",'red'), lty=1, cex=0.8)

# people recovered in day 5 and 10
out_SIR %>% filter(time %in% c(5,10)) %>% select(time,n_R)
```
Day 5: 35784 people \
Day 10: 76370 people

#### Different solver testing
Gain experience points by experimenting
with different solvers to see if they have an effect.
```{r}

## reasonable numerical solution with euler
eu_R <- as.data.frame(euler(x, t, SIR_R, parameters))
eu_S <- as.data.frame(euler(y, t, SIR_S, parameters)) ; eu_S$S[is.nan(eu_S$S)]<-0  # replace NaN with 0
eu_I <- left_join(eu_R,eu_S, by=c('time'='time')) %>% mutate(I=1-R-S)

# plot SIR graphs (time step is adjusted to 0.5)
# plot R
plot(eu_R$t, eu_R$R, type='l', lty=3, col = "blue", cex = 0.5,  ylim=c(0,1), xlab='time', ylab='fraction of total population', main = 'Plotting S, I, R curves')
# plot S
points(eu_S$t, eu_S$S, type='l',lty=3, col = "black", cex = 0.5)
# plot I
points(eu_I$t, eu_I$I, type='l',lty=3, col = "red", cex = 0.5)

# Rk4 original plot
# plot R
points(out_R$t, out_R$R, type='l', col = "blue", cex = 0.5)
# plot S
points(out_S$t, out_S$S, type='l', col = "black", cex = 0.5)
# plot I
points(out_I$t, out_I$I, type='l', col = "red", cex = 0.5)
# legend
legend(25,0.9,legend=c("R_euler", "S_euler","I_euler","R_rk4", "S_rk4","I_rk4" ),
       col=c("blue", "black",'red'), lty=c(3,3,3, 1,1,1), cex=0.8)

```
Euler method and Runge-Kutta 4th order method has some difference in their estimates. 
Runge-Kutta 4th order method is proved to be more accurate than the Euler method.





