---
title: "Homework1"
author: "KiseokUChicago"
date: "2021-01-22"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=9, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

### Homework1
## Coding assignment for ECEV 42900
Professor: **Sarah Cobey** **Greg Dwyer** 
Student: **Kiseok Lee**  

### Exercise 6. The Ricker Model and Bifurcation diagram. Repeat the analysis of the fixed point for the Ricker Model:

This is the provided script to make bifurcation plot for the logistic map
```{r}
## Bifurcation.R 
## Stefano Allesina sallesina@uchicago.edu allesinalab.uchicago.edu
## Code for "Theoretical Ecology"


## This function returns the values of the min and max 
peaks <- function(x) {
  if (min(x)==max(x)) return(min(x)) ## Does not oscillate 
  l <- length(x) 
  xm1 <- c(x[-1], x[l])
  xp1 <- c(x[1], x[-l]) 
  z<-x[x > xm1 & x > xp1 | x < xm1 & x < xp1] 
  if (length(z)==0) return(min(x)) ## It has not converged yet 
  return (z)
} 

## This function creates a simulation of the logistic map 
LogisticMap<-function(N0,r,TimeSteps){
  Results<-rep(0,TimeSteps) 
  Results[1]<-N0 
  for (j in 2:TimeSteps){
    Results[j]<-r*Results[j-1]*(1-Results[j-1])
  } 
  return(Results)
} 

## Plot the Diagram 
plot(0,0, xlim=c(0,4), ylim=c(-0.05,1.05),type="n", xlab="r", ylab="X") 
for (r in seq(0.001,4,0.005)) { # These are the initial and final values for r
  out <- LogisticMap(0.5,r,2500) # Initial conditions 
  l <- length(out) %/% 10 # use only the last 250 steps 
  out <- out[(9*l):(10*l)] 
  p <- peaks(out) 
  l <- length(out) 
  points(rep(r, length(p)), p, pch=".")
}


```


Make new function for Ricker model
```{r}
## This function creates a simulation of the logistic map 
RickerModel<-function(N0,r,TimeSteps,b){
  Results<-rep(0,TimeSteps) 
  Results[1]<-N0 
  for (j in 2:TimeSteps){
    Results[j]<-r*Results[j-1]*exp(-b*Results[j-1])
  } 
  return(Results)
} 

## Plot the Diagram with b fixated (4 different values)

par(mfrow=c(2,2))
b_vec <- c(0.5, 2, 4, 8) 
r_end <- 20

for (b in b_vec){
  plot(0,0, xlim=c(0,r_end), ylim=c(-0.05,1.05),type="n", xlab="r", ylab="X")
  for (r in seq(0.001,r_end,0.005)) { # These are the initial and final values for r
  out <- RickerModel(0.5,r,2500,b) # Initial conditions 
  l <- length(out) %/% 10 # use only the last 250 steps 
  out <- out[(9*l):(10*l)] 
  p <- peaks(out) 
  l <- length(out) 
  points(rep(r, length(p)), p, pch=".")
}
}

```
Ricker model is: N(t+1)=rN(t)e^(-bN(t)) \
From the top left to bottom right, b value is 0.5, 2, 4(bottom left), 8(bottom right).
We can see first bifurcations around r=1 in all 4 graphs. 


### Exercise 7. Numerical integration.

Integrate the exponential growth and the logistic growth using as initial condition t = 0,N(0) = 0.25 until t = 10. \
Plot the error at time t = 10 as a function of the step-size (using delta t = (0.05, 0.1, 0.2)).

This is the provided script to make Euler method approximations on dN/dt = rN (exponential growth)
```{r}
library(ggplot2)
library(ggthemes)

ExponentialGrowth<-function(x,lambda){
   return (lambda*x)
}

Euler<-function(x, D, FUN,lambda){
   return(x+FUN(x,lambda)*D)
}

ExponentialSolution<-function(N0,t,lambda){
   return (N0*exp(t*lambda))
}
# dev.off()
N0<-0.25
lambda<-0.3
plot(c(0,0),col="0",xlim=c(0,10),ylim=c(0,(N0*exp(lambda*10)+1)), xlab='t', ylab='x')
for (Deltat in c(0.05,0.1,0.2)){
   PointsToEstimate<-seq(0,10,by=Deltat)
   Iterations<-length(PointsToEstimate)-1
   Approx<-rep(0,Iterations+1)
   RealSol<-rep(0,Iterations+1)
   Approx[1]<-RealSol[1]<-N0
   for (i in 1:Iterations){
      RealSol[i+1]<-ExponentialSolution(N0,Deltat*i,lambda)		
      Approx[i+1]<-Euler(Approx[i],Deltat,ExponentialGrowth,lambda)		
   }
   if (Deltat==0.05) mycol="red"
   if (Deltat==0.1) mycol="blue"
   if (Deltat==0.2) mycol="green"
   points(RealSol~PointsToEstimate,col="black",type="l")
   points(Approx~PointsToEstimate,col=mycol,type="l")
}

# error at time t=10 as a function of step-size
get_error <- function(Deltat=0.05, t=10){
  PointsToEstimate<-seq(0,10,by=Deltat)
  Iterations<-length(PointsToEstimate)-1
  Approx<-rep(0,Iterations+1)
  RealSol<-rep(0,Iterations+1)
  Approx[1]<-RealSol[1]<-N0
  for (i in 1:Iterations){
    RealSol[i+1]<-ExponentialSolution(N0,Deltat*i,lambda)		
    Approx[i+1]<-Euler(Approx[i],Deltat,ExponentialGrowth,lambda)		
  }
  return (RealSol[t] - Approx[t])
}

get_error(0.05, 10)
df_1 <- data.frame(deltat=c(0.05, 0.1, 0.2), error=NA)
df_1$error <- sapply(df_1$deltat,get_error)
df_1

ggplot(df_1, aes(x=deltat, y=error)) + geom_point() + theme_hc() +
    ggtitle("Error at t=10 in different stepsizes(0.05, 0.1, 0.2)")

```


Add the function to logistic growth
I increased the error time t=10 to t=30 in order to better see the plateau.
```{r}
LogisticGrowth<-function(x,lambda){
   return (lambda*x*(1-x))
}

LogisticSolution<-function(N0,t,lambda){
   return ( (N0*exp(t*lambda)) / ( 1+N0*(exp(t*lambda)-1) ) )
}

# dev.off()
N0<-0.25
lambda<-0.3
plot(c(0,0),col="0",xlim=c(0,30),ylim=c(0,(N0*exp(lambda*2)+1)), xlab='t', ylab='x')
for (Deltat in c(0.05,0.1,0.2)){
   PointsToEstimate<-seq(0,30,by=Deltat)
   Iterations<-length(PointsToEstimate)-1
   Approx<-rep(0,Iterations+1)
   RealSol<-rep(0,Iterations+1)
   Approx[1]<-RealSol[1]<-N0
   for (i in 1:Iterations){
      RealSol[i+1]<-LogisticSolution(N0,Deltat*i,lambda)		
      Approx[i+1]<-Euler(Approx[i],Deltat,LogisticGrowth,lambda)		
   }
   if (Deltat==0.05) mycol="red"
   if (Deltat==0.1) mycol="blue"
   if (Deltat==0.2) mycol="green"
   points(RealSol~PointsToEstimate,col="black",type="l")
   points(Approx~PointsToEstimate,col=mycol,type="l")
}

```
The Euler's estimation of logistic growth model works well. The red/blue/green line almost aligns with the black line, which is the real solution of the logistic growth model.

### Exercise 8. More numerical integration!

Using the code below as a template, write functions that integrate the exponential, logistic, Allee, Levins, and theta-logistic ODE models we saw in this chapter using deSolve. \
```{r}
## ExpGrowth.R 
## Stefano Allesina sallesina@uchicago.edu allesinalab.uchicago.edu
## Code for "Theoretical Ecology"

require(deSolve) 
## package for integrating numerically ODEs see 
##http://cran.r-project.org/web/packages/deSolve/
```

Plot the dynamics over time. Specify which parameter values you use, and label everything. \

#### (1) exponential ODE model
```{r}
## This function takes a step in time
ExponentialGrowth<-function(t, state, parameters) { 
  X <- state[1] ## the first element is the density of X at time t
  r <- parameters[1] ## the first parameter is the growth rate
  dX<-r*X 
  return(list(dX)) ## for some reason, you have to return a list
}

## This function runs the model and produces the trajectory
RunExponentialGrowth <- function(MaxTime=10, GrowthRate=0.1, InitialX=1.0, title='Exponential'){
  times <- seq(0, MaxTime, by = 0.01)
  parameters <- c(r=GrowthRate)
  state <- c(X=InitialX)
  out <- ode(y = state, times = times, func = ExponentialGrowth, parms = parameters)
  print(plot(out[,2]~out[,1],type="l",xlab="time",ylab="Density of X", main=title))
  # return(out)
}

RunExponentialGrowth()

```

#### (2) logistic model
```{r}
## This function takes a step in time
LogisticGrowth<-function(t, state, parameters) { 
  X <- state[1] ## the first element is the density of X at time t
  r <- parameters[1] ## the first parameter is the growth rate
  K <- parameters[2]
  dX<-r*X*(1-(X/K))
  return(list(dX)) ## for some reason, you have to return a list
}

## This function runs the model and produces the trajectory
RunLogisticGrowth <- function(MaxTime=100, GrowthRate=0.1, InitialX=1.0, Carrying_capa=15, title='Logistic'){
  times <- seq(0, MaxTime, by = 0.01)
  parameters <- c(r=GrowthRate, K=Carrying_capa)
  state <- c(X=InitialX)
  out <- ode(y = state, times = times, func = LogisticGrowth, parms = parameters)
  print(plot(out[,2]~out[,1],type="l",xlab="time",ylab="Density of X", main=title))
  # return(out)
}

RunLogisticGrowth()
```

#### (3) Allee effect model
```{r}
## This function takes a step in time
AlleeGrowth<-function(t, state, parameters) { 
  X <- state[1] ## the first element is the density of X at time t
  u <- parameters[1] ## the first parameter is the growth rate
  c <- parameters[2]
  v <- parameters[3]
  dX<- X*( (u*X/(v+X))-c*X )
  return(list(dX)) ## for some reason, you have to return a list
  # u = 2, c = 0.1, and v = 5.
}

## This function runs the model and produces the trajectory
RunAlleeGrowth <- function(MaxTime=10, InitialX=2, u=2, c=0.1, v=5, title='Allee effect'){
  times <- seq(0, MaxTime, by = 0.01)
  parameters <- c(u=u, c=c, v=v)
  state <- c(X=InitialX)
  out <- ode(y = state, times = times, func = AlleeGrowth, parms = parameters)
  print(plot(out[,2]~out[,1],type="l",xlab="time",ylab="Density of X", main=title))
  # return(out)
}

RunAlleeGrowth()

```

#### (4) Levins model
```{r}

## This function takes a step in time
LevinsGrowth<-function(t, state, parameters) { 
  X <- state[1] ## the first element is the density of X at time t
  c <- parameters[1] ## the first parameter is the growth rate
  dX<- c*X*(1-X) - exp(1)*X
  return(list(dX)) ## for some reason, you have to return a list
}

## This function runs the model and produces the trajectory
RunLevinsGrowth <- function(MaxTime=10, c=1, InitialX=1.0, title='Levins'){
  times <- seq(0, MaxTime, by = 0.01)
  parameters <- c(c=c)
  state <- c(X=InitialX)
  out <- ode(y = state, times = times, func = LevinsGrowth, parms = parameters)
  print(plot(out[,2]~out[,1],type="l",xlab="time",ylab="Density of X", main=title))
  # return(out)
}

RunLevinsGrowth()

```

#### (5) theta-logistic ODE model
```{r}
## This function takes a step in time
ThetaLogisticGrowth<-function(t, state, parameters) { 
  X <- state[1] ## the first element is the density of X at time t
  r <- parameters[1] ## the first parameter is the growth rate
  K <- parameters[2]
  theta <- parameters[3]
  dX<- r*X*( 1 - (X/K)^theta)
  return(list(dX)) ## for some reason, you have to return a list
}

## This function runs the model and produces the trajectory
RunThetaLogisticGrowth <- function(MaxTime=30, r=.25, K=10, theta = 1, InitialX=1.0, title='ThetaLogistic'){
  times <- seq(0, MaxTime, by = 0.01)
  parameters <- c(r=r, K=K, theta=theta)
  state <- c(X=InitialX)
  out <- ode(y = state, times = times, func = ThetaLogisticGrowth, parms = parameters)
  print(plot(out[,2]~out[,1],type="l",xlab="time",ylab="Density of X", main=title))
  # return(out)
}

# plot for 3 different theta value as Exercise 3
par(mfrow=c(3,1))
for (th in c(0.5, 1, 1.5)){
  RunThetaLogisticGrowth(theta=th)
}

```


