---
title: "Homework8"
author: "KiseokUChicago"
date: "2021-03-11"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=9, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

### Homework8 (Bayesian)
## Coding assignment for ECEV 42900
Professor: **Sarah Cobey**, **Greg Dwyer** \
Student: **Kiseok Lee**  


```{r}
#A function for plotting error bars
ErrBars<-function(x,y,SE){
	NumPts = length(x);
	for(Pts in 1:NumPts){
		xvals = c(x[Pts],x[Pts]);
		Upper = y[Pts]+SE[Pts];
		Lower = y[Pts]-SE[Pts];
		yvals = c(Upper,Lower);
		lines(xvals,yvals);
	}
}


#Reading in the data
Data = read.csv("data/VirusData6.csv");

######### Downhill Simplex to Generate Proposal Means

# Examining the data
colnames(Data)
head(Data)

# combination function
comb = function(n, x) {
  factorial(n) / factorial(n-x) / factorial(x)
}

#The log-likelihood function with heterogeneity
logLHoodHtg<-function(par){

	betaBar = par[1];
	V  = par[2];

	timeT = 7.0;
	qoft = (1+ betaBar*V*(Data$Density)*timeT)^(-1/V) ; #1. Type in the model that describes the probability of surviving to time T
	n0 = Data$Total; #n0 is the total number of insects in each treatment
	N = Data$Uninfected; #N is the number uninfected
	logLHood = log(comb(n0,N))+N*log(qoft)+(n0-N)*log(1-qoft); #2. Type in the log-likelihood function for the pure death model
	return(-sum(logLHood));  #We are looking for the negative sum of the log likelihoods across experimental units, because optim is a minimizer

}

# test the function
par=c(0.1,0.1)
logLHoodHtg(par) # success

#Finding the best-fit parameter values
HtgOptOut = optim(par=c(0.1,0.1),fn=logLHoodHtg);
BestBeta = HtgOptOut$par[1]; 
BestVee = HtgOptOut$par[2];

Tmts = unique(Data$Density); #Figure out what the treatments were
AvgTransmn = tapply(-log(Data$Uninfected/Data$Total),Data$Density,mean); #Use tapply to calculate the mean transmission rateat each treatment
SDTransmn = tapply(-log(Data$Uninfected/Data$Total),Data$Density,sd); #Use tapply to calculate the standard deviation of the transmission rate

par(mfrow=c(1,1));
par(mai=c(1,1,1,1));
plot(Tmts,AvgTransmn,ylim=c(0,2),xlim=c(0,max(Tmts)),xlab="Cadavers per branch",ylab="Transmission"); #Plot the model prediction against the data


NumReps = nrow(Data)/length(Tmts); #Count how many replicates there were, for the standard-error calculation
ErrBars(x=Tmts,y=AvgTransmn,SDTransmn/sqrt(NumReps-1)); #Plot the error bars - standard error is sd/sqrt(n-1)


Cadavers = seq(from=0,to=4,by=0.1);
time = 7;
BestHtgModel = (1/BestVee)*log(1 + BestBeta*time*BestVee*Cadavers);
lines(Cadavers,BestHtgModel,col="RED",lwd=2.5);




############## MCMC: running the chains #####################

maxitn = 2e5; maxchain = 5;
LH = c(); MaxPars = 2;

ThinStop = 1000;
ThinStor = maxitn/ThinStop;
StorPars = array(dim=c(maxchain,ThinStor,MaxPars));

PropMean = c(); PropSD = c(); Prior = c()
#2. Fix the proposals.  Do the means on a log scale.
PropMean[1] = log(BestBeta); 
PropMean[2] = log(BestVee); 
PropSD[2] = 2.0;  
PropSD[1] = 2.0; 


Prior[1] = 1000.0;
Prior[2] = 1000.0;
ScaleSD = 1.2;

for(chain in 1:maxchain){
	CurrentPars = c(); #the current values of the parameters
	NewPars = c(); #the proposed values of the parameters
	for(i in 1:MaxPars){ #starting values for the parameters
		CurrentPars[i] = exp(rnorm(1,mean=PropMean[i],sd = ScaleSD*PropSD[i]));
	}
	#Calculating the posterior of the initial parameter set
	OldLH = -1.0*logLHoodHtg(par=CurrentPars); #-1 because logLHoodHtg returns -log lhood, and we want log lhood
	OldPost = OldLH;
	for(i2 in 1:MaxPars){
		OldPost = OldPost + log(dunif(x=(CurrentPars[i2]),max=Prior[i2]));
	}
	#parHtg = c(exp(BestBeta),exp(BestVee));
	#logLHoodHtg(parHtg);

	AcceptCount = 0;
	itnStor = 1;
	for(itn in 1:maxitn){
	  
	
		if(itn==1){
			ScaleSD = 1.2;
		}else{
			ScaleSD = 1.0;
		}
		for(i in 1:MaxPars){
			# print(paste0('i',i))
			for(i2 in 1:MaxPars){
				 NewPars[i2] = CurrentPars[i2];
			}
			#3.  Propose a new parameter value, and calculate its posterior.
		  # i=1
		  # i=2
			NewPars[i] = exp(rnorm(1,mean=PropMean[i],sd = ScaleSD*PropSD[i]));
			NewLH = -1.0*logLHoodHtg(par=NewPars); 
			NewPost = NewLH;
			for(i2 in 1:MaxPars){
			 	NewPost = NewPost + log(dunif(x=(NewPars[i2]),max=Prior[i2])) ;
			}

			#Here is the adjustment for the proposal for the current parameter value
			OldPropAdj = log(dnorm(x=log(CurrentPars[i]),mean=PropMean[i],sd=ScaleSD*PropSD[i]));
		
			#4.  Now do the adjustment for the proposal, based on the new parameter value
			NewPropAdj = log(dnorm(x=log(NewPars[i]),mean=PropMean[i],sd=ScaleSD*PropSD[i]))
			
			# print(paste0('NewPost',NewPost))
			# print(paste0('NewPropAdj',NewPropAdj))
			
			# print(paste0('OldPost',OldPost))
			# print(paste0('OldPropAdj',OldPropAdj))

			if(itn>1){
			 	#6. Write out the acceptance criterion, and decide whether to accept or not 
				#6a. First, calculate the acceptance criterion
				Criterion = exp(NewPost-NewPropAdj-(OldPost-OldPropAdj));
				# print(paste0('Criterion ',Criterion))

				#6b. Second, draw a U(0,1) random variate.
			 	temp = runif(1, min=0, max=1);
			 	# print(paste0('temp ',temp))

				#6c. If statement for whether or not to accept
			 	if(Criterion >= temp){
			 	  # print(paste0('Chain: ',chain,' / itn: ',itn," / ",Criterion >= temp, " / NewPars: ",NewPars))
					AcceptCount = AcceptCount + 1;
					OldPost = NewPost;
					OldLH = NewLH;
					#7. Acceptance means setting current parameters equal to proposed parameters
					for(i2 in 1:MaxPars){
				 		 CurrentPars[i2] = NewPars[i2]  ;
					}#for i2
			 	} #if Criterion
			} #itn>1
		} #for i in 1:MaxPars

	    #We only want to keep some of the iterations
	    if((itn %% ThinStop)==0){
    		cat("chain:",chain,"itn:",itn,"itnStor:",itnStor);
    		for(i2 in 1:MaxPars){
    			StorPars[chain,itnStor,i2] = CurrentPars[i2];
    			cat(" i2:",i2,"StorPars:",StorPars[chain,itnStor,i2]);
		}
		itnStor = itnStor + 1;
		cat("\n"); #line = readline();
		
	    } #Thin

					
	} #itn

} #chain

########### MCMC diagnostics: did the algorithm converge? #############
require(coda);
AllPars = c();
itnStor = itnStor - 1;
HalfWay = round(0.5*itnStor);
x1 = mcmc(StorPars[1,HalfWay:itnStor,]);
x2 = mcmc(StorPars[2,HalfWay:itnStor,]);
x3 = mcmc(StorPars[3,HalfWay:itnStor,]);
x4 = mcmc(StorPars[4,HalfWay:itnStor,]);
x5 = mcmc(StorPars[5,HalfWay:itnStor,]);
xAll = mcmc.list(x1,x2,x3,x4,x5);

#Trace plots
par(ask="TRUE");
par(mai=c(0.5,0.5,0.5,0.5));
plot(xAll)
gelmanOut = gelman.diag(xAll)
print(gelmanOut);

#Autocorrelations
par(mfrow=c(5,2));
par(ask="TRUE");
par(mai=c(0.225,0.25,0.225,0.25));
#par(mai=c(0.5,0.5,0.5,0.5));
title(main="")
acf(x1[,1],xlab="",ylab=""); abline(h=-0.1,col="RED"); abline(h=0.1,col="RED");
acf(x1[,2],xlab="",ylab=""); abline(h=-0.1,col="RED"); abline(h=0.1,col="RED");
acf(x2[,1],xlab="",ylab=""); abline(h=-0.1,col="RED"); abline(h=0.1,col="RED");
acf(x2[,2],xlab="",ylab=""); abline(h=-0.1,col="RED"); abline(h=0.1,col="RED");
acf(x3[,1],xlab="",ylab=""); abline(h=-0.1,col="RED"); abline(h=0.1,col="RED");
acf(x3[,2],xlab="",ylab=""); abline(h=-0.1,col="RED"); abline(h=0.1,col="RED");
acf(x4[,1],xlab="",ylab=""); abline(h=-0.1,col="RED"); abline(h=0.1,col="RED");
acf(x4[,2],xlab="",ylab=""); abline(h=-0.1,col="RED"); abline(h=0.1,col="RED");
acf(x5[,1],xlab="",ylab=""); abline(h=-0.1,col="RED"); abline(h=0.1,col="RED");
acf(x5[,2],xlab="",ylab=""); abline(h=-0.1,col="RED"); abline(h=0.1,col="RED");

	

############### Decision-making: should we spray the virus? #################
AllData = c();
for(chain in 1:maxchain){
	AllData = rbind(AllData,StorPars[chain,1:ThinStor,]);
}


Burnout<-function(nu,mu,V,S0,I0,i){ #This function is a straight line

	return (1 - i - (1 + nu*V/mu*(i*S0+I0))^(-V));

}

S0 = 0.1; I0 = 0.01;
UpperPctl = round(0.25*ThinStor);
LowerPctl = round(0.75*ThinStor);
S0Range = seq(from=0.0,to=10.0,by=0.1); #the range of values over which we will calculate the root
S0Dim = length(S0Range);
ParamDim = nrow(AllData);
#FractInf = array(dim=c(S0Dim,ParamDim));
FractInf = c();
FIPlotUpper = c(); FIPlotLower = c(); FIMedian = c();
i = 1;
for(S0 in S0Range){
	for(j in 1:ThinStor){
	
		z = uniroot(f=Burnout,interval=c(0.0,1.0),nu=AllData[j,1],mu=0.41,V=AllData[j,2],S0=S0,I0=I0);
		FractInf[j] = z$root;
		#cat("S0:",S0,"j:",j,"FractInf:",FractInf[j],"\n");

		if(1){
			Ordering = order(FractInf,decreasing=TRUE); #figuring out the order of the parameters
	 		OrderedFractInf = cbind(FractInf)[Ordering,] #Carrying out the ordering...
			FIPlotUpper[i] = OrderedFractInf[UpperPctl];
			FIPlotLower[i] = OrderedFractInf[LowerPctl];
			FIMedian[i] = median(FractInf);
		}

	}
	#line = readline();
	i = i + 1;
}

par(ask="FALSE");



k1 = 0.0;
k2 = 100.0;
#k2 = 0.05;

par(mfrow=c(1,1));

MedianCost = k1 + k2*(1-FIMedian)*S0Range;
LowerCost = k1 + k2*(1-FIPlotUpper)*S0Range;
UpperCost = k1 + k2*(1-FIPlotLower)*S0Range;
#plot(S0Range,(1-FIMedian)*S0Range*k1*exp(k2*Defoln),type="l"); #exponential
par(mfrow=c(1,1));
Upperylim = max(UpperCost);
Lowerylim = min(LowerCost);
par(mai=c(1,1,1,1));
plot(S0Range,MedianCost,type="l",ylim=c(Lowerylim,Upperylim),ylab="Cost",xlab="Insect Density");  #linear
lines(S0Range,UpperCost,type="l",lwd=2.5);
lines(S0Range,LowerCost,type="l",lwd=2.5);



```

Thought questions
1. The forest service personnel that we work with have sometimes been dismayed when they spray high density populations, but get almost the same infection rate as in populations at the same density that are unsprayed. What does epidemiological theory say could be the explanation? \

According to figure 4, the higher the insect density (starting from density > 2), the lower the cost of not spraying. It means that in order to get the maximum effect of spraying we would need to spray when the insect density if low (around 1). In other words, spraying is less effective in high insect density (6~) compared to lower density (0.5~2). \

2. At low insect densities, the cost of not spraying increases with increasing density,  because of the defoliation caused by the insect. Eventually, however, the cost of not spraying falls very sharply with increasing host density, leading to a peak in the cost curve. The curve at first increases with increasing insect density, but why does it then fall very sharply? Be sure to invoke the disease-density threshold in your answer. \

For density dependent diseases, there exists a threshold at which the transmission of the pathogen becomes inefficient. This is called the disease-density threshold. The threshold host density of a species refers to the minimum concentration of individuals necessary to sustain a given disease within a population (Wikipedia). Therefore, the reason the cost of not spraying plummets after the threshold is because the pathogen has already reached and passed the point where it cannot sustain itself due to its transmission being ineffective.

3. Suppose that the cost of spraying is 40$/hectare (vertical axis units on the figure are in dollars/hectare). What is the highest density at which you would recommend spraying? Also, note that the horizontal axis scale is insects /m2. If the pest manager then says that costs above 60/hectare are absolutely unacceptable, what is the highest density at which you would recommend spraying? Be sure to invoke the 75th percentile line in your answer.

The cost of spraying has to be less than cost of not spraying(depicted in graph) in order to justify the act of spraying. Therefore, if we use the 75th percentile of cost of not spraying as our worst case scenario, we could get the highest insect density where the cost of not spraying(75th percentile) is lower or equal to 60/hectare. 
```{r}
plot(S0Range,MedianCost,type="l",ylim=c(Lowerylim,Upperylim),ylab="Cost",xlab="Insect Density");  #linear
lines(S0Range,UpperCost,type="l",lwd=2.5);
lines(S0Range,LowerCost,type="l",lwd=2.5);
abline(h=60, col='red')

S0Range[UpperCost < 60]

abline(v=0.65, col='blue')
```

There are two points where the 75th percentile graph intersects with cost = 60$. However, it would be recommended to spray before the insect density gets higher and already damage the ecosystem. That density is 0.65insects/m^2 (blue line). 










