---
title: "Homework5"
author: "KiseokUChicago"
date: "2021-02-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=9, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

### Homework5
## Coding assignment for ECEV 42900
Professor: **Sarah Cobey**, **Greg Dwyer** \
Student: **Kiseok Lee**  

### Assignment 1: Environmental Stochasticity
Your goal in the homework is to reproduce fig. 1. Set the reproductive rate lambda = 2, the standard
deviation of the log of lambda sigma = 0.5. What this means is that you will have to draw normally
distributed random variates with mean 2 and standard deviation 0.5. Iterate the model for 10
time steps, as in the upper panel of the figure. In the lower panel of the figure, I have plotted
a histogram of the final population size (logN10) for 10,000 realizations of the model. Upload
your code to the Canvas site.

### Given code
```{r}
#Declaring parameter values
lambda = 2.0;
Rlzns = 10000;
StDev = 0.5;
TimeSteps = 10;
InitN = 1.0;

#Declaring that N and Stor are vectors
N <- numeric();
Stor <- numeric();

#When we plot N vs t, it is helpful to have t already holding the values of time
t = c(0:(TimeSteps));

#We are going to plot two graphs, the next line sets that up for us
par(mfrow=c(2,1));


for(j in 1:Rlzns){ #This is a for loop, so that we can plot many realizations
	N[1] = InitN #1. Change this line so that N[1] gets the initial value of N
	for(i in 2:(TimeSteps+1)){ #This for loop iterates the model over the number of time steps.
	  eps <- rnorm(1,mean=0,sd=StDev)
		N[i] = lambda*exp(eps)*N[i-1] #2. Change this line so that N[i] is the new value of the population size
	}
	Stor[j] = log(N[i]) #3. Change this line so that Store gets the LAST value of N

	if(j==1){ #For the first realization, we use "plot"
	    # print(N)
			plot(t,log(N),type="l",xaxs="i",ylim=c(-3,12),xlim=c(0,10),xlab="Time (generations)",ylab="log Pop. Size");
	}else{ #For subsequent realizations, we use "lines", to add lines to the original plot
	    # print(N)
			lines(t,log(N));	
	}
}

#Next line is a function that plots a straight line, with intercept "a" and slope "b"
abline(a=log(InitN),b=log(lambda),col="RED",lwd=4); #4. Calculate the average for abline
	
#Plot a histogram of the population sizes in "Stor"
hist(Stor,freq=FALSE,breaks=200,right=TRUE,xlab="Final population size",ylab="Frequency")
#Create a vector "x"...
x = seq(from=-20,to=20,by=0.1);
TheoryAvg = log(InitN) + TimeSteps*log(lambda) #5.  Insert the theoretical expectation expression here 
TheoryVar = TimeSteps*(StDev^2) #6. Insert the theoretical variance expression here
#Calculate the probability distribution function of a normal distribution using the theoretical expectation and variance
Dist = dnorm(x,mean=TheoryAvg,sd=TheoryVar^(0.5));
#Plot the distribution function
lines(x,Dist,lwd=1,col="RED");
```

### 5.6.1
Suppose that for given species E(lambda) > 1, sigma >> 1. \
Explain why the model implies that the long-term prospects for the survival of this species are not very good, even though the long-term expected size is large.

```{r}
#Declaring parameter values, Change the sigma from 0.5 to 5
lambda = 2.0;
Rlzns = 30;
StDev = 5;
TimeSteps = 10;
InitN = 1.0;

#Declaring that N and Stor are vectors
N <- numeric();
Stor <- numeric();

#When we plot N vs t, it is helpful to have t already holding the values of time
t = c(0:(TimeSteps));

#We are going to plot two graphs, the next line sets that up for us
par(mfrow=c(2,1));


for(j in 1:Rlzns){ #This is a for loop, so that we can plot many realizations
	N[1] = InitN #1. Change this line so that N[1] gets the initial value of N
	for(i in 2:(TimeSteps+1)){ #This for loop iterates the model over the number of time steps.
	  eps <- rnorm(1,mean=0,sd=StDev)
		N[i] = lambda*exp(eps)*N[i-1] #2. Change this line so that N[i] is the new value of the population size
	}
	Stor[j] = log(N[i]) #3. Change this line so that Store gets the LAST value of N

	if(j==1){ #For the first realization, we use "plot"
	    # print(N)
			plot(t,log(N),type="l",xaxs="i",ylim=c(-3,12),xlim=c(0,10),xlab="Time (generations)",ylab="log Pop. Size");
	}else{ #For subsequent realizations, we use "lines", to add lines to the original plot
	    # print(N)
			lines(t,log(N));	
	}
}

#Next line is a function that plots a straight line, with intercept "a" and slope "b"
abline(a=log(InitN),b=log(lambda),col="RED",lwd=4); #4. Calculate the average for abline
	
#Plot a histogram of the population sizes in "Stor"
hist(Stor,freq=FALSE,breaks=200,right=TRUE,xlab="Final population size",ylab="Frequency")
#Create a vector "x"...
x = seq(from=-20,to=20,by=0.1);
TheoryAvg = log(InitN) + TimeSteps*log(lambda) #5.  Insert the theoretical expectation expression here 
TheoryVar = TimeSteps*(StDev^2) #6. Insert the theoretical variance expression here
#Calculate the probability distribution function of a normal distribution using the theoretical expectation and variance
Dist = dnorm(x,mean=TheoryAvg,sd=TheoryVar^(0.5));
#Plot the distribution function
lines(x,Dist,lwd=1,col="RED");

```

When the variation of lambda(t) is large, hence the large sigma, population can go to extinction as shown in the time series graph. When lambda becomes less than 1 a few times, the population plummets.
