---
title: "Homework3"
author: "KiseokUChicago"
date: "2021-02-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=9, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

### Homework3
## Coding assignment for ECEV 42900
Professor: **Sarah Cobey**, **Greg Dwyer** \
Student: **Kiseok Lee**  

### Exercise 2. Community stability
(1) For S = 250, sigma = 0.3, and C between 0.01 and 0.5, compute the
probability of observing a stable random and a stable predator-prey matrix. \
(2) To estimate the value, simulate at least 100 matrices
for each level of C. Draw a profile of C vs. the probability
of stability for the two cases. \
(3) Which critical value of C would May’s theorem produce? Are the simulations consistent with
this value?

### Given code
```{r}
## Community_matrices.R
## Stefano Allesina and Sarah Cobey 
## Code for Theoretical Ecology

MayMatrix<-function(S,C,sigma){
  ## This matrix determines the connections
  A<-matrix(runif(S*S),S,S)
  ## Contains the values for the connections
  B<-matrix(rnorm(S*S,0.0,sigma),S,S)
  A<-(A<= C)*1 # A matrix contains 1 when A[i,j] <= C
  M<-A*B
  diag(M)<- -1
  return(M)
}

PPMatrix<-function(S,C,sigma){
  ## Determine the signs for the connections
  MyS<-sign(rnorm(S*(S-1)/2))
  A<-matrix(0,S,S)
  A[upper.tri(A,diag=F)]<-MyS
  D<-matrix(runif(S*S),S,S)
  D<-(D <= C)*1
  A<-D*A
  A<- A-t(A)
  ## Contains the  values for the connections
  B<-matrix(abs(rnorm(S*S,0.0,sigma)),S,S)
  M<-A*B
  diag(M)<- -1
  return(M)
}

```

### Getting the critical value C (probability) in May's matrix 
For S = 250, sigma = 0.3, and C between 0.01 and 0.5, compute the
probability of observing a stable random matrix.
To estimate the value, simulate at least 100 matrices
for each level of C. Draw a profile of C vs. the probability
of stability for the two cases.

```{r}
library(schoolmath)

# Testing with May matrix
# MayMatrix(S=250, 0.01, sigma=0.3)
# Getting eigenvalues 
eig_values <- eigen(MayMatrix(S=250, 0.01, sigma=0.3), only.values = T)$values
# Real part of the complex number (eigen value)
Re(eig_values)
# are all real part of the eigenvalues negative?
all(is.negative(Re(eig_values))) # all negative eigenvalues indicate stable


# Getting probabilities for May's matrix
df_c_may <- data.frame(C=NA, prob=NA)

for (C in seq(0.01, 0.5, by=0.01)){ # 0.01, 0.02, 0.03,...., 0.50
  counter=0
  for (i in c(1:100)){
    # get the eigenvalues of the community matrix
    eig_values <- eigen(MayMatrix(S=250, C, sigma=0.3), only.values = T)$values
    # real part of the complex number
    Re(eig_values)
    # negative eigenvalue
    if(all(is.negative(Re(eig_values)))){ # all negative eigenvalues indicate stable community
      counter = counter +1
    } 
  }
  df_c_may <- rbind(df_c_may, c(C, counter/100))
}

df_c_may1 <- df_c_may[-1,]

library(ggplot2)
p1 <- ggplot(df_c_may1,aes(x = C,y = prob)) + geom_line(col="magenta") + geom_point(color='maroon') + theme_bw() +
  labs(title="Probability of matrix being stable (May's matrix) \n", x="\n C (probability of drawing 1)",y="probability of having stable community \n")
p1
```

Threshold of C for May's case is 0.04. After 0.04, probability of the community matrix being stable plummets. \

```{r}
# May's critical value
# C = ((1/sigma)^2 ) / S
(1/0.3)^2 / 250

```

May's theorem produces critical C value of 0.044. This is very close to value of 0.04 from the simulation.


### Getting the critical value C (probability) in Prey-predator's matrix 
For S = 250, sigma = 0.3, and C between 0.01 and 0.5, compute the
probability of observing a stable predator-prey matrix.
To estimate the value, simulate at least 100 matrices
for each level of C. Draw a profile of C vs. the probability
of stability for the two cases.

```{r}
# Getting probabilities for Prey-predator's matrix
df_c_pp <- data.frame(C=NA, prob=NA)

for (C in seq(0.01, 0.5, by=0.01)){
  counter=0
  for (i in c(1:100)){
    # get the eigenvalues of the community matrix
    eig_values <- eigen(PPMatrix(S=250, C, sigma=0.3), only.values = T)$values
    # real part of the complex number
    Re(eig_values)
    # negative eigenvalue
    if(all(is.negative(Re(eig_values)))){ # all negative eigenvalues indicate stable community
      counter = counter +1
    } 
  }
  df_c_pp <- rbind(df_c_pp, c(C, counter/100))
}

df_c_pp1 <- df_c_pp[-1,]

library(ggplot2)
p2 <- ggplot(df_c_pp1,aes(x = C,y = prob)) + geom_line(col="magenta") + geom_point(color='maroon') + theme_bw() +
  labs(title="Probability of matrix being stable (Prey-predator's matrix) \n", x="\n C (probability of drawing 1)",y="probability of having stable community \n")
p2

```

Threshold of C for Allesina & Pascual's case is between 0.25 to 0.35. Gradually after 0.25, probability of the community matrix being stable plummets.

### (2) Examples of population dynamics

Lotka-Volterra model: two species x and y
dx/dt = lambda * x - epsilon * x * y
dy/dt = eta * x * y - delta * y

```{r}
library(phaseR)

lotkaVolterra <- function(t, y, parameters) {
x <- y[1]
y <- y[2]
lambda <- parameters[1]
epsilon <- parameters[2]
eta <- parameters[3]
delta <- parameters[4]
dy <- numeric(2)
dy[1] <- lambda*x - epsilon*x*y
dy[2] <- eta*x*y - delta*y
list(dy)
}

# When (lambda, epsilon, eta, delta) = (2, 1, 3, 2)
lotkaVolterra.flowField <- flowField(lotkaVolterra, xlim = c(0, 5), ylim = c(0, 10),parameters = c(2, 1, 3, 2), points = 19, add = FALSE)
grid()
lotkaVolterra.nullclines <- nullclines(lotkaVolterra, xlim = c(-1, 5), ylim = c(-1, 10),parameters = c(2, 1, 3, 2), points = 500)
y0 <- matrix(c(1, 2, 2, 2, 3, 4), ncol = 2, nrow = 3, byrow = TRUE)
lotkaVolterra.trajectory <- trajectory(lotkaVolterra, y0 = y0, tlim =c(0,10), parameters = c(2, 1, 3, 2), col = rep("black", 5))



```


```{r}

model <- function(t, y, parameters) {
x <- y[1]
y <- y[2]
lambda <- parameters[1]
epsilon <- parameters[2]
eta <- parameters[3]
delta <- parameters[4]
dy <- numeric(2)
dy[1] <- lambda*x + epsilon*y
dy[2] <- eta*x + delta*y
list(dy)
}

# When (lambda, epsilon, eta, delta) = (2, 1, 3, 2)
model.flowField <- flowField(model, xlim = c(0, 10), ylim = c(0, 10),parameters = c(-1, -1, 1, -3), points = 19, add = FALSE)
grid()
model.nullclines <- nullclines(model, xlim = c(-1, 5), ylim = c(-1, 10),parameters = c(-1, -1, 1, -3), points = 500)
y0 <- matrix(c(1, 2, 2, 2, 3, 4), ncol = 2, nrow = 3, byrow = TRUE)
model.trajectory <- trajectory(model, y0 = y0, tlim =c(0,10), parameters = c(-1, -1, 1, -3), col = rep("black", 5))


```


